---
title: "03_Manual_Filter"
output: html_notebook
author: "Matthew J. Brooks"
date_created: "October 28th, 2022"
---


This analysis filters the 06_200310 Seurat object of doublet cells using manual technique and gets rid of low-feature rods to generate the final filtered dataset.

### Import Seurat list objects 
```{r, message=FALSE}
library(Seurat)

# Get run info data
seu <- readRDS("../data/interim/seu_merged_221028.rds") # This is used from the output of 01_Norm_Annotation
# seu.filt <- readRDS("../data/interim/seu_filt.06_200310_221021.rds") # This is used to continue the analysis after the manual filter step

```


### Create directories
```{r}

expt_nm <- "06_200310_dX_df_dX"
dir_export <- file.path("../data/processed", expt_nm, "03_Manual_Filter")

dir.create(dir_export, recursive = T)

```

<!-- ### Re-call the cell type annotation using Sanes/NNRL markers -->
<!-- ```{r} -->
<!-- library(tidyverse) -->
<!-- source("../src/data/sctype_ann_v2.R") -->

<!-- ## Sanes cell-type -->
<!-- gs_sanes_file <- "/Volumes/Accelsior_8M2/scRNAseq/scRNA_aging/221001/data/external/Cell_type_MB.xlsx" -->

<!-- raw_gs_sanes <- readxl::read_excel(gs_sanes_file) -->

<!-- gs_sanes_sub <- list() -->
<!-- for (i in unique(raw_gs_sanes$sub_cell_type)){ -->
<!--   tmp <- raw_gs_sanes %>%  -->
<!--     dplyr::filter(sub_cell_type == i) %>%  -->
<!--     dplyr::select(gene) %>%  -->
<!--     as.vector() %>%  -->
<!--     unlist(use.names = F) %>%  -->
<!--     toupper() -->
<!--   gs_sanes_sub[[i]] <- tmp -->
<!-- } -->

<!-- sctypes = c("Astrocytes", "Immune cells", "Retinal pigment epithelial cells") -->


<!-- seu.06_200310 <- sctype_ann(seu.06_200310, gs_pos = gs_sanes_sub,  -->
<!--                        sctypes = c("Astrocytes", "Immune cells", "Retinal pigment epithelial cells")) -->

<!-- # Add sub cell type annotation to metadata -->
<!-- seu.06_200310 <- AddMetaData(seu.06_200310, metadata = seu.06_200310$CellType, col.name = "SubCellType") -->

<!-- # Get subcell to cell type named vector for recoding -->
<!-- celltype_ls <- raw_gs_sanes$cell_type -->
<!-- names(celltype_ls) <- raw_gs_sanes$sub_cell_type -->
<!-- celltype_ls <- celltype_ls[unique(names(celltype_ls))] -->
<!-- celltype_ls <- celltype_ls[grep("one", celltype_ls)] -->

<!-- # recode the subcell types as cell types -->
<!-- sub_cell <- seu.06_200310$SubCellType -->
<!-- args_sub <- c(list(sub_cell), celltype_ls) -->
<!-- sub_cell_replace <- do.call(recode, args_sub) -->

<!-- # Add cell type annotation to meta data -->
<!-- seu.06_200310 <- AddMetaData(seu.06_200310, metadata = sub_cell_replace, col.name = "CellType") -->

<!-- # Make plots -->
<!-- a <- DimPlot(seu.06_200310, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'SubCellType') -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_SubCellTypeAnn.png")), width = 16, height = 12) -->

<!-- a <- DimPlot(seu.06_200310, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'CellType') -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_CellTypeAnn.png")), width = 16, height = 12) -->




<!-- ``` -->


## Get gene markers for seu res=0.8
```{r}
seu.celltype <- seu

# Change Idents to vell type
Idents(seu.celltype) <- seu.celltype$CellType

# Find cell type markers
seu.cellmarkers <- FindAllMarkers(seu.celltype, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write.table(seu.cellmarkers, file = file.path(dir_export, paste0(expt_nm, "_celltype_markers_res08.tsv")), quote = F, row.names = F, sep = "\t")

# seu.cellmarkers %>%
#     group_by(cluster) %>%
#     top_n(n = 10, wt = avg_log2FC) -> top10
# 
# t10 <- DoHeatmap(seu, features = top10$gene) + NoLegend()
# cowplot::ggsave2(plot = t10, file = "../data/processed/06_200310_celltype_Top10_res08.png", width = 24, height = 32)


```


## Increase cluster resolution and find marker genes for 
```{r}
seu.tmp <- seu

# Res 6
seu.tmp <- FindClusters(seu.tmp, resolution = 6)
u <- DimPlot(seu.tmp, reduction = "umap", label = T, pt.size = 0.1) + NoLegend()
cowplot::ggsave2(plot = u, file = file.path(dir_export, paste0(expt_nm, "_Dim_cluster_res6.png")), width = 12, height = 12)

seu.markers.tmp <- FindAllMarkers(seu.tmp, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
write.table(seu.markers.tmp, file = file.path(dir_export, paste0(expt_nm, "_cluster_markers_res6.tsv")),
            quote = F, row.names = F, sep = "\t")

# seu.markers.tmp %>%
#     group_by(cluster) %>%
#     top_n(n = 10, wt = avg_log2FC) -> top10
# 
# t10 <- DoHeatmap(seu.tmp, features = top10$gene) + NoLegend()
# cowplot::ggsave2(plot = t10, file = "../data/processed/06_200310_BusDub_Top10_res6.png", width = 24, height = 32)
```


## Investigate top 3 cell type marker genes on clusters
```{r}
library(tidyverse)

df.tmp <- seu.tmp@meta.data %>% group_by(seurat_clusters) %>% dplyr::select(CellType) %>% table() %>% as.tibble() %>% dplyr::filter(n > 1) %>% dplyr::rename(cells=n) %>% group_by(seurat_clusters) %>% arrange(desc(cells)) %>% slice_head(n=1) %>% arrange(CellType)



seu.tmp$seurat_clusters <- factor(x = seu.tmp$seurat_clusters, levels = rev(df.tmp$seurat_clusters))
# seu.tmp@meta.data$seurat_clusters <- factor(x = seu.tmp@meta.data$seurat_clusters, levels = df.tmp$seurat_clusters)
Idents(seu.tmp) <- seu.tmp$seurat_clusters

# df.tmp <- seu.tmp@meta.data %>% group_by(seurat_clusters) %>% dplyr::select(CellType) %>% table() %>% as.tibble() %>% dplyr::filter(n > 1) %>% group_by(seurat_clusters) %>% slice_max(desc(n)) %>% arrange(CellType)
# 
# seu.tmp$seurat_clusters <- factor(x = seu.tmp$seurat_clusters, levels = rev(df.tmp$seurat_clusters))
# # seu.tmp@meta.data$seurat_clusters <- factor(x = seu.tmp@meta.data$seurat_clusters, levels = df.tmp$seurat_clusters)
# Idents(seu.tmp) <- seu.tmp$seurat_clusters

# goi <- c("Rho", "Gnat1", "Pde6b", "Sag",
#          "Opn1sw", "Opn1mw", "Gnat2", "Pde6c", "Arr3",
#          "Apoe", "Glul", "Rlbp1", "Clu")
cellmark_mod <- seu.cellmarkers
cellmark_mod$cluster <- factor(cellmark_mod$cluster, levels = sort(levels(seu.cellmarkers$cluster)))

cellmark_filt <- cellmark_mod %>% 
  dplyr::filter(pct.2 < 0.25) %>%
  arrange(cluster) %>% 
  group_by(cluster) %>% 
  slice_max(order_by = avg_log2FC, n=3)
  
col.side <- scales::hue_pal()(length(unique(seu.tmp$CellType)))
names(col.side) <- sort(unique(seu.tmp$CellType))
col.clst <- col.side[df.tmp$CellType]
clst.cumsum <- names(col.clst) %>% table() %>% as.vector() %>% rev() %>% cumsum()

goi <- cellmark_filt$gene
names(goi) <- rep(col.side, each = 3)[1:length(goi)]

  
q <- VlnPlot(seu.tmp, goi, stack = T) + NoLegend()

q <- q + 
  scale_y_discrete(labels = paste(rev(names(col.clst)), levels(Idents(seu.tmp)), sep = "_")) +
  theme(axis.text.y = element_text(colour = rev(col.clst)),
        axis.text.x = element_text(colour = names(goi)),
               plot.background = element_rect(fill = "white"),
               panel.background = element_rect(fill = "white")) +
  geom_hline(yintercept = c(clst.cumsum + 0.5), 
             linetype = "dashed", color = "darkgrey") 

cowplot::ggsave2(plot = q, filename = file.path(dir_export, paste0(expt_nm, "_Viol_Top3filtCellMarker_res6.png")), height = 24, width = 48)
```


## Filter out the unwanted clusters with multiple canonical markers and features < 600
```{r}
# source("../src/data/sctype_ann.R")

# Remove clusters with multiple canonical markers or unknown cell type
seu.filt <- subset(seu.tmp, idents = c(78, 58, 81, 61, 101), invert = T)

# Remove cells with less than 600 Features/cell
feat_lt_600 <- seu.filt$nFeature_RNA < 600
seu.filt <- AddMetaData(seu.filt, metadata = feat_lt_600, "feat_lt_600")

# seu.final <- subset(seu.filt, subset = "feat_lt_600", invert = T)
seu.final <- seu.filt[,!(seu.filt@meta.data$feat_lt_600)]

```


### Renormalize
```{r}
source("../src/data/opt_PC.R")

pc <- opt_PC(seu.final)

# Re-normalize and UMAP
seu.final <- SCTransform(seu.final, assay = "decontXcounts") %>% 
  RunPCA() %>% 
  FindNeighbors(dims = 1:20) %>%
  RunUMAP(dims = 1:20) %>% 
  FindClusters()

```


### Re-annotate cell type
```{r}

source("../src/data/sctype_ann_v2.R")

## Sanes cell-type
gs_sanes_file <- "/Volumes/Accelsior_8M2/scRNAseq/scRNA_aging/221001/data/external/Cell_type_MB.xlsx"

raw_gs_sanes <- readxl::read_excel(gs_sanes_file)

gs_sanes_sub <- list()
for (i in unique(raw_gs_sanes$sub_cell_type)){
  tmp <- raw_gs_sanes %>% 
    dplyr::filter(sub_cell_type == i) %>% 
    dplyr::select(gene) %>% 
    as.vector() %>% 
    unlist(use.names = F) %>% 
    toupper()
  gs_sanes_sub[[i]] <- tmp
}

sctypes = c("Astrocytes", "Immune cells", "Retinal pigment epithelial cells")


seu.final<- sctype_ann(seu.final, gs_pos = gs_sanes_sub, 
                       sctypes = c("Astrocytes", "Immune cells", "Retinal pigment epithelial cells"))

# Add sub cell type annotation to metadata
seu.final <- AddMetaData(seu.final, metadata = seu.final$CellType, col.name = "SubCellType")

# Get subcell to cell type named vector for recoding
celltype_ls <- raw_gs_sanes$cell_type
names(celltype_ls) <- raw_gs_sanes$sub_cell_type
celltype_ls <- celltype_ls[unique(names(celltype_ls))]
celltype_ls <- celltype_ls[grep("one", celltype_ls)]

# recode the subcell types as cell types
sub_cell <- seu.final$SubCellType
args_sub <- c(list(sub_cell), celltype_ls)
sub_cell_replace <- do.call(recode, args_sub)

# Add cell type annotation to meta data
seu.final <- AddMetaData(seu.final, metadata = sub_cell_replace, col.name = "CellType")

```


### Make final plots
```{r}

# Change idents to cell type
# Idents(seu.final) <- seu.filt$CellType

# Make the UMAP plots
u <- DimPlot(seu.final, reduction = "umap", label = T, pt.size = 0.1) + NoLegend() + ggtitle("Filtered")
cowplot::ggsave2(plot = u, file = file.path(dir_export, paste0(expt_nm, "_Dim_final.png")), width = 12, height = 12)

a <- DimPlot(seu.final, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'CellType')
cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_CellTypeAnn_final.png")), width = 16, height = 12)

b <- DimPlot(seu.final, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'SubCellType')
cowplot::ggsave2(plot = b, filename = file.path(dir_export, paste0(expt_nm, "_Dim_SubCellTypeAnn_final.png")), width = 16, height = 12)
```


## VAriable genes for Soumitra Oct 26th, 2022
```{r}

var_gene <- row.names(seu.final@assays$SCT@scale.data)

write.csv(var_gene, file = paste0(dir_export, "/06-200310_final_SCT_varGenes.csv"))

```


<!-- ## Test plot -->
<!-- ```{r} -->

<!-- a <- DimPlot(seu.filt, reduction = "umap", split.by = "age", group.by = "tissue") -->

<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_indi_age_filt.png")), width = 48, height = 12) -->
<!-- ``` -->



## Get gene markers for seu res=0.8
```{r}

# Change Idents to cell type
Idents(seu.filt) <- seu.filt$seurat_clusters

# Find cell type markers
seu.clustermarkers <- FindAllMarkers(seu.filt, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write.table(seu.clustermarkers, file = file.path(dir_export, paste0(expt_nm, "_cluster_markers_filt_res08.tsv")), quote = F, row.names = F, sep = "\t")

# seu.cellmarkers %>%
#     group_by(cluster) %>%
#     top_n(n = 10, wt = avg_log2FC) -> top10
# 
# t10 <- DoHeatmap(seu, features = top10$gene) + NoLegend()
# cowplot::ggsave2(plot = t10, file = "../data/processed/06_200310_celltype_Top10_res08.png", width = 24, height = 32)

Idents(seu.filt) <- seu.filt$CellType
```


<!-- ## Show Features less than 600, filter and re-cluster  -->
<!-- ```{r} -->
<!-- library(tidyverse) -->
<!-- feat_lt_600 <- seu.filt$nFeature_RNA < 600 -->
<!-- seu.filt <- AddMetaData(seu.filt, metadata = feat_lt_600, "feat_lt_600") -->

<!-- # seu.final <- subset(seu.filt, subset = "feat_lt_600", invert = T) -->
<!-- seu.final <- seu.filt[,!(seu.filt@meta.data$feat_lt_600)] -->

<!-- seu.final <- SCTransform(seu.final, assay = "decontXcounts") %>%  -->
<!--   RunPCA() %>%  -->
<!--   FindNeighbors(dims = 1:20) %>% -->
<!--   RunUMAP(dims = 1:20) %>%  -->
<!--   FindClusters() -->

<!-- # # Re-annotate -->
<!-- # seu.final <- sctype_ann(seu.final, gs_pos = gs_sanes_sub,  -->
<!-- #                        sctypes = c("Astrocytes", "Immune cells", "Retinal pigment epithelial cells")) -->

<!-- # # Add sub cell type annotation to metadata -->
<!-- # seu.final <- AddMetaData(seu.final, metadata = seu.final$CellType, col.name = "SubCellType") -->
<!-- #  -->
<!-- # # Get subcell to cell type named vector for recoding -->
<!-- # celltype_ls <- raw_gs_sanes$cell_type -->
<!-- # names(celltype_ls) <- raw_gs_sanes$sub_cell_type -->
<!-- # celltype_ls <- celltype_ls[unique(names(celltype_ls))] -->
<!-- # celltype_ls <- celltype_ls[grep("one", celltype_ls)] -->
<!-- #  -->
<!-- # # recode the subcell types as cell types -->
<!-- # sub_cell <- seu.final$SubCellType -->
<!-- # args_sub <- c(list(sub_cell), celltype_ls) -->
<!-- # sub_cell_replace <- do.call(recode, args_sub) -->
<!-- #  -->
<!-- # # Add cell type annotation to meta data -->
<!-- # seu.final <- AddMetaData(seu.final, metadata = sub_cell_replace, col.name = "CellType") -->


<!-- # Plots -->
<!-- a <- FeaturePlot(seu.filt, features = "feat_lt_600") -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Feat_filt_lt600.png")), width = 16, height = 12) -->

<!-- a <- DimPlot(seu.final, reduction = "umap", label = TRUE, repel = TRUE) -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_clusters_final.png")), width = 16, height = 12) -->

<!-- a <- DimPlot(seu.final, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'CellType') -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_CellTypeAnn_final.png")), width = 16, height = 12) -->

<!-- a <- DimPlot(seu.final, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'SubCellType') -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_SubCellTypeAnn_final.png")), width = 16, height = 12) -->


<!-- ``` -->


## Cell Marker plots
```{r}

dir_export2 <- file.path("../data/processed", expt_nm, "03_Manual_Filter", "CellAnnGOIplots")

dir.create(dir_export2)


for (i in 1:dim(raw_gs_sanes)[1]){

  tmp_goi <- raw_gs_sanes$gene[i]
  tmp_outname <- paste(gsub(" ", "", raw_gs_sanes$sub_cell_type[i]), tmp_goi, sep = "_")
  
  if (tmp_goi %in% row.names(seu.final@assays$RNA@counts)){
    x <- FeaturePlot(seu.final, features = tmp_goi) + ggtitle(tmp_goi)
  cowplot::ggsave2(plot = x, filename = file.path(dir_export2, paste0(expt_nm, "_Feat_", tmp_outname, ".png")), width = 16, height = 12)
  }
  
}

```

## Determine optimal PCs for PCA analysis
https://hbctraining.github.io/scRNA-seq/lessons/elbow_plot_metric.html

```{r}
# Plot the elbow plot
ElbowPlot(object = seu.final, 
          ndims = 50)

# Determine percent of variation associated with each PC
pct <- seu.final[["pca"]]@stdev / sum(seu.final[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]

co1

# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

# last point where change of % of variation is more than 0.1%.
co2

# Minimum of the two calculation
pcs <- min(co1, co2)

pcs

# Create a dataframe with values
plot_df <- data.frame(pct = pct, 
           cumu = cumu, 
           rank = 1:length(pct))

# Elbow plot to visualize 
  ggplot(plot_df, aes(cumu, pct, label = rank, color = rank > pcs)) + 
  geom_text() + 
  geom_vline(xintercept = 90, color = "grey") + 
  geom_hline(yintercept = min(pct[pct > 5]), color = "grey") +
  theme_bw()



```



<!-- ```{r} -->
<!-- seu <- seu.filt -->

<!-- testnm <- "pca15" -->

<!-- seu <- RunPCA(seu) %>%  -->
<!--   FindNeighbors(dims = 1:15) %>% -->
<!--   RunUMAP(dims = 1:15) %>%  -->
<!--   FindClusters() -->


<!-- a <- DimPlot(seu, reduction = "umap", label = TRUE, repel = TRUE) -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_clusters_final_", testnm, ".png")), width = 16, height = 12) -->

<!-- a <- DimPlot(seu, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'CellType') -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_CellTypeAnn_final_", testnm, ".png")), width = 16, height = 12) -->

<!-- a <- DimPlot(seu, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'SubCellType') -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_SubCellTypeAnn_final_", testnm, ".png")), width = 16, height = 12) -->




<!-- ``` -->

## Make barplot for numbers of cells filtered out
```{r}

cell_filt_og <- read.csv("../data/processed/06-200310_Cell_filt_count.tsv", sep = "\t")
colnames(cell_filt_og) <- c("State", "Cells", "Source")
cell_filt <- cell_filt_og %>% dplyr::filter(!(State %in% c("orig_cells", "filt_cells")))
cell_interim <- cell_filt_og %>% dplyr::filter(State == "filt_cells")                         

cell_final <- table(seu.filt$orig.ident) %>% as.data.frame() %>% mutate(State = "final_cells")
cell_final <- cell_final[c(3,2,1)]
colnames(cell_final) <- c("State", "Cells", "Source")

cell_manual <- data.frame(State = "manual_filt",
                          Cells = cell_interim$Cells - cell_final$Cells,
                          Source = cell_final$Source)

cell_counts <- rbind(cell_filt, cell_manual, cell_final)
cell_counts$State <- factor(cell_counts$State, levels = levels(factor(cell_counts$State))[c(3,1,4,5,2)])
cell_counts$Source <- factor(cell_counts$Source, levels = unique(cell_counts$Source))

b <- cell_counts %>% 
  ggplot(aes(x = Source, fill = State, y=Cells)) +
  geom_bar(position = "stack", stat = "identity") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  annotate("text", x=seq(1,8), y = cell_final$Cells-500, label = cell_final$Cells)

cowplot::ggsave2(plot = b, filename = "../data/processed/06-200310_Filt_Bar_cellcounts.pdf", height = 3, width = 6)
```

<!-- # Rgr expression -->
<!-- ```{r} -->

<!-- mueller_cells <- seu.final$CellType == "Mueller cell" -->
<!-- rgr_cells <- sum(GetAssayData(seu.final, slot = "data")["Rgr", mueller_cells]>0) -->
<!-- rgr_percet <- rgr_cells / length(mueller_cells[mueller_cells == T]) -->

<!-- seu.tmp <- seu.final -->
<!-- Idents(seu.tmp) <- seu.tmp$CellType -->
<!-- a <- FeaturePlot(seu.tmp, features = "Rgr", label = T, repel = T, pt.size = 0.5) -->
<!-- cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_Rgr.png")), width = 16, height = 12) -->
<!-- ``` -->


# Export seurat object
```{r}

# saveRDS(object = seu.filt, file = "../data/interim/seu_filt.06_200310_221021.rds")
saveRDS(object = seu.final, file = "../data/interim/seu_final.06_200310_22109.rds")

```