---
title: "01_Norm_Annotation"
output: html_notebook
author: "Matthew J. Brooks"
date_created: "October 28th, 2022"
---


This analysis merges the Seurat list object, performs decontX, then normalization, and assigns cell type annotation.

### Import Seurat list objects, download Cell type file, helper scripts, and create directories 
```{r, message=FALSE}
library(Seurat)

proj_nm <- "experiment"
expt_nm <- proj_nm

script_dir <- "https://raw.githubusercontent.com/MBrooks313/scRNA-seq/main"
dest_tools_dir <- "../src/tools"

################
# Get Seurat list object from 00_Data_import
seu_list <- readRDS("../data/interim/seu_list_00_import.rds")


################
# Download cell type file
download.file(url = file.path(script_dir, "Retina_Cell_type_MB.xlsx"), 
              destfile = file.path(dest_tools_dir, "Retina_Cell_type_MB.xlsx"), 
              method = "curl")


################
# Download helper scripts
download.file(url = file.path(script_dir, "sctype_ann.R"), 
              destfile = file.path(dest_tools_dir, "sctype_ann.R"), 
              method = "curl")


################
# Create directories

dir_export <- file.path("../data/processed", expt_nm, "01_Norm_Annotation")
dir.create(dir_export, recursive = T)

```

### Merge Seurat list into one Seurat object
```{r, message=FALSE}

################
# Merge data

seu <- merge(x = seu_list[[1]], seu_list[2:length(seu_list)], 
                       add.cell.ids = names(seu_list), 
                       project = proj_nm)

```


### Add meta features to Seurat object
```{r}

################
# Add data to meta.data
# age <- gsub("(.+)_.+", "\\1", seu@meta.data$orig.ident)
# tissue <- gsub(".+_(.+)", "\\1", seu@meta.data$orig.ident)

# seu <- AddMetaData(seu, age, col.name = "age")
# seu <- AddMetaData(seu, tissue, col.name = "tissue")

# Adjust factors
# seu@meta.data$orig.ident <- factor(seu@meta.data$orig.ident, 
                                    # levels = unique(seu@meta.data$orig.ident))

# seu@meta.data$age <- factor(seu@meta.data$age, 
                                # levels = unique(seu@meta.data$age))

# seu@meta.data$tissue <- factor(seu@meta.data$tissue, 
                                # levels = unique(seu@meta.data$tissue))

```


### Run decontX on merged data
```{r}
source(file.path(dest_tools_dir, "decontX.R"))

seu <- decX(seu)

```

### Normalize and cluster the Seurat object
```{r, message=FALSE}

library(sctransform)
library(kableExtra)

seu <- SCTransform(seu, assay = "decontXcounts") %>% 
    RunPCA() %>%
  FindNeighbors(dims = 1:30) %>% 
  RunUMAP(dims = 1:30) %>% 
  FindClusters()

```



### Add Ccell type annotation
```{r, message=FALSE}
library(tidyverse)

source(file.path(dest_tools_dir, "sctype_ann.R"))

##################
## Sanes cell-type

gs_sanes_file <- file.path(dest_tools_dir, "Retina_Cell_type_MB.xlsx")

raw_gs_sanes <- readxl::read_excel(gs_sanes_file)

gs_sanes_sub <- list()
for (i in unique(raw_gs_sanes$sub_cell_type)){
  tmp <- raw_gs_sanes %>% 
    dplyr::filter(sub_cell_type == i) %>% 
    dplyr::select(gene) %>% 
    as.vector() %>% 
    unlist(use.names = F) %>% 
    toupper()
  gs_sanes_sub[[i]] <- tmp
}

sctypes = c("Astrocytes", "Immune cells", "Retinal pigment epithelial cells")


################
# Perform cell type annotation at sub cell type level
seu <- sctype_ann(seu, gs_pos = gs_sanes_sub, 
                  sctypes = c("Astrocytes", "Immune cells", "Retinal pigment epithelial cells"),
                  assay_data = "SCT")

# Add sub cell type annotation to metadata
seu <- AddMetaData(seu, metadata = seu$CellType, col.name = "SubCellType")


################
# Recode the sub cell type to cell type and add to meta

# Get subcell to cell type named vector for recoding
celltype_ls <- raw_gs_sanes$cell_type
names(celltype_ls) <- raw_gs_sanes$sub_cell_type
celltype_ls <- celltype_ls[unique(names(celltype_ls))]
celltype_ls <- celltype_ls[grep("one", celltype_ls)]

# recode the subcell types as cell types
sub_cell <- seu$SubCellType
args_sub <- c(list(sub_cell), celltype_ls)
sub_cell_replace <- do.call(recode, args_sub)

# Add cell type annotation to meta data
seu <- AddMetaData(seu, metadata = sub_cell_replace, col.name = "CellType")


################
# Make plots
a <- DimPlot(seu, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'SubCellType')
cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_SubCellTypeAnn_merged.png")), width = 16, height = 12)

a <- DimPlot(seu, reduction = "umap", label = TRUE, repel = TRUE, group.by = 'CellType')
cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_CellTypeAnn_merged.png")), width = 16, height = 12)

a <- DimPlot(seu, reduction = "umap", label = TRUE, repel = TRUE)
cowplot::ggsave2(plot = a, filename = file.path(dir_export, paste0(expt_nm, "_Dim_cluster_merged.png")), width = 16, height = 12)

```


### Get cell type gene markers for seu
```{r}
seu.celltype <- seu

# Change Idents to vell type
Idents(seu.celltype) <- seu.celltype$CellType

# Find cell type markers
seu.cellmarkers <- FindAllMarkers(seu.celltype, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write.table(seu.cellmarkers, file = file.path(dir_export, paste0(expt_nm, "_celltype_markers_merged.tsv")), quote = F, row.names = F, sep = "\t")

# seu.cellmarkers %>%
#     group_by(cluster) %>%
#     top_n(n = 10, wt = avg_log2FC) -> top10
# 
# t10 <- DoHeatmap(seu, features = top10$gene) + NoLegend()
# cowplot::ggsave2(plot = t10, file = "../data/processed/06_200310_celltype_Top10_res08.png", width = 24, height = 32)

```


## Get cluster gene markers for seu
```{r}

# Find cluster markers
seu.clustermarkers <- FindAllMarkers(seu, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

write.table(seu.clustermarkers, file = file.path(dir_export, paste0(expt_nm, "_cluster_markers_merged.tsv")), quote = F, row.names = F, sep = "\t")

# seu.cellmarkers %>%
#     group_by(cluster) %>%
#     top_n(n = 10, wt = avg_log2FC) -> top10
# 
# t10 <- DoHeatmap(seu, features = top10$gene) + NoLegend()
# cowplot::ggsave2(plot = t10, file = "../data/processed/06_200310_celltype_Top10_res08.png", width = 24, height = 32)


```


### Export Seurat object
```{r}

saveRDS(object = seu, file = "../data/interim/seu_merged_01_norm.rds")

```